---
title: "Using Descriptive Statistics to Analyze Data in R: Coursera Guided Project"
author: "Leopoldine Mirtil"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### DATA SOURCE

The data is from the “Using Descriptive Statistics to Analyze Data in R” guided project course offered through Coursera.com and taught by instructor Nikunj Maheshwari. 

### OBJECTIVE

Guided project to calculate descriptive statistical metrics on a data set and create a data quality report file.

### Task 1 - Load and View Data

#### Set Directory 

```{r}
setwd("C:/Users/Leopoldine/Desktop/Mine/Coding Projects & Portfolio/Guided Projects/R Guided Project/DescriptiveStatisticsInR")
```


#### Load Library

```{r, message=FALSE}
library(tidyverse)
library(rmarkdown)
```


#### Import and View Data

```{r}
data <- read.csv("data.csv")

head(data)
```


### Task 2 - Calculate the Measure of Frequency Metrics

#### Number of Observations in Column

```{r}
length(data$Transmission.Type)
```


#### Number of Unique Categories
 
```{r}
length(unique(data$Transmission.Type))   
```


#### Calculate Frequency Distribution of each Unique Category in a Column 

```{r}
table(data$Transmission.Type)
```


#### Store Output into Variable

```{r}
freq <- sort(table(data$Transmission.Type), decreasing = TRUE) 

freq
```


### Task 3 - Calculate the Measure of Central Tendency Metrics 

#### Mean
  
```{r}
mean(data$Engine.HP)   

mean(data$Engine.HP, na.rm=TRUE)  # 'na.rm=TRUE' removes missing values 
```


#### Median

```{r}
median(data$Engine.HP, na.rm=TRUE) 
```


#### Mode

```{r}
uniqueValues <- unique(data$Engine.HP) #finds all unique values in column and store it in variable

uniqueValues [which.max(tabulate(match(data$Engine.HP, uniqueValues)))]  
```


### Task 4 - Calculate the Measure of Dispersion Metrics 

#### Minimum

```{r}
min(data$Engine.HP, na.rm= TRUE)  
```


#### Maximum

```{r}
max(data$Engine.HP, na.rm= TRUE)
```


#### Range

```{r}
range(data$Engine.HP, na.rm= TRUE) # gives mean and max values
```


#### Variance

```{r}
var(data$Engine.HP, na.rm= TRUE)  #measures how much each value is varying/deviating from mean
```


#### Standard Variance

```{r}
sd(data$Engine.HP, na.rm= TRUE) #measures average deviation of each value of mean
```


### Task 5 - Use R’s in-built Functions for Additional Data Quality Metrics

#### Find Data Type of Column

```{r}
class(data$Engine.HP) 
```

#### Calculate Number of Missing Values

```{r}
sum(is.na(data$Transmission.Type)) 

sum(is.na(data$Number.of.Doors))
```


### Task 6 - Create Custom R Function to Calculate Descriptive Statistics on any given Dataset 

#### Count Total Values in each Column
 
```{r}
apply(data, MARGIN=2, length)
```


#### Find Minimum Value of each Column

```{r}
sapply(data, function(x) min(x, na.rm=TRUE)) 
```


#### Revise Import Formula

```{r}
data <- read.csv("data.csv", stringsAsFactors = FALSE) 
```

#### Create Dataframe showing all Descriptive 

```{r}
quality_data <- function(df=NULL){
  if(is.null(df)) print("Please pass a non-empty data frame")
  summary_table <- do.call(data.frame, 
                           list(
                             Min = sapply(df, function(x) min(x, na.rm=TRUE)),
                             Max = sapply(df, function(x) max(x, na.rm=TRUE)),
                             Mean = sapply(df, function(x) mean(x, na.rm=TRUE)),
                             SD = sapply(df, function(x) sd(x, na.rm=TRUE)),
                             Total = apply(df, 2, length),
                             NULLS = sapply(df, function(x) sum(is.na(x))),
                             Unique = sapply(df, function(x) length(unique(x))),
                             dataType = sapply(df, class)
                           ))
  nums <- vapply(summary_table, is.numeric, FUN.VALUE = logical(1))
  summary_table[,nums] <- round(summary_table[, nums], digits = 3)
  return(summary_table)
}
```


### Task 7 - Export Results of Descriptive Statistics to a Data Quality Report File 

#### Use New Function to View Summary Statistics

```{r, warning=FALSE, message=FALSE, error=FALSE}
df_quality <- quality_data(data)

df_quality 
```

#### Convert Row Names into Separate Columns

```{r, warning=FALSE, message=FALSE}
df_quality <- cbind(Columns=rownames(df_quality), 
                    data.frame(df_quality, row.names = NULL))  


df_quality 
```

#### Export File

```{r}
write.csv(df_quality, "Data Quality Report.csv", row.names = FALSE)
```

#### Export File without Overwriting

```{r}
write.csv(df_quality, paste("Data Quality Report",
          format(Sys.time(), "%d-%m-%Y-%H%M%S"),
          ".csv"), 
          row.names = FALSE)
```
